---
title: "ROC p-values"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# AUC comparison: Boostrap method
## AUC difference significance testing
 
   * Computation details:
   a) boot.n (2000) bootstrap replicates are drawn from the data. If boot.stratified = TRUE, each replicate contains exactly the same number of controls and cases than the original sample
   
   b) for each bootstrap replicate, the AUC of the two ROC curves are computed and the difference is used:
   $$  D =  \frac{AUC_1-AUC_2}{std(AUC_1-AUC_2)} \sim Z$$
     where $std$ is the standard deviation of the bootstrap differences and AUC1 and AUC2 the AUC of the two (bootstrap replicate) ROC curves.
   
   c) Z approximately follows a normal distribution, one or two-tailed p-values can be calculated accordingly.

## For exp1: 
```{r}
library(pROC)
library(readr)
setwd("Z:/Cristina/Section3/paper_notes_section3_MODIFIED")
exp1_pooled_pred_train <- read_csv("datasets/exp1b_pooled_pred_train.csv")
exp1_pooled_pred_val <- read_csv("datasets/exp1b_pooled_pred_val.csv")
exp1_pred_test <- read_csv("datasets/exp1b_pred_test.csv")

# Create 6 objects: Tr_exp1, ciobj_Tr_exp1, Val_exp1, ciobj_val_exp1, Te_exp1, ciobj_Te_exp1
# plot ROC curve
################
library(pROC)
par(mfrow = c(1, 3))
par(cex = 0.6)
par(mar = c(3, 3, 0, 0), oma = c(0.51, 0.51, 0.51, 0.51))
n=12
colors = rainbow(n, s = 1, v = 1, start = 0, end = max(1, n - 1)/n, alpha = 1)
# plot
Tr_exp1 <- plot.roc(exp1_pooled_pred_train$labels, exp1_pooled_pred_train$probC, col=colors[1], lty=1)
ciobj_Tr_exp1 <- ci.se(Tr_exp1, specificities=seq(0, 1, 0.05)) 
plot(ciobj_Tr_exp1, type="shape", col="grey") # plot as a grey shape
par(new=TRUE)
plot.roc(exp1_pooled_pred_train$labels, exp1_pooled_pred_train$probC, col=colors[1], lty=1, main="ROC for cvTrain")
legend("bottomright", 
       legend = c(paste0("cvTrain: AUC=", format(Tr_exp1$auc,digits=3, format="f"))), 
       col = colors[1],lwd = 2, lty = c(1))

# plot
Val_exp1 <- plot.roc(exp1_pooled_pred_val$labels, exp1_pooled_pred_val$probC, col=colors[1], lty=1)
ciobj_val_exp1 <- ci.se(Val_exp1, specificities=seq(0, 1, 0.05)) 
plot(ciobj_val_exp1, type="shape", col="grey") # plot as a blue shape
par(new=TRUE)
plot.roc(exp1_pooled_pred_val$labels, exp1_pooled_pred_val$probC, col=colors[10], lty=1, main="ROC for cvValidation")
legend("bottomright", 
       legend = c(paste0("cvValidation: AUC=", format(Val_exp1$auc, digits=3, format="f"))), 
       col = colors[10],lwd = 2, lty = c(1))
# plot
Te_exp1 <- plot.roc(exp1_pred_test$labels, exp1_pred_test$probC, col=colors[8], lty=1)
ciobj_Te_exp1 <- ci.se(Te_exp1, specificities=seq(0, 1, 0.05)) 
plot(ciobj_Te_exp1, type="shape", col="grey") # plot as a grey shape
par(new=TRUE)
plot.roc(exp1_pred_test$labels, exp1_pred_test$probC, col=colors[8], lty=1, main="ROC for held-out Test")
legend("bottomright", 
       legend = c(paste0("Test: AUC=", formatC(Te_exp1$auc,digits=2, format="f"))), 
       col = colors[8],lwd = 2, lty = c(1))

```

* Do we significantly overtrain?

```{r exp1-overtrain, echo=TRUE}
roc.test(Tr_exp1, Val_exp1, method="bootstrap",boot.stratified=TRUE, alternative="greater")

roc.test(Tr_exp1, Te_exp1, method="bootstrap",boot.stratified=TRUE, alternative="greater")

roc.test(Val_exp1, Te_exp1, method="bootstrap",boot.stratified=TRUE, alternative="two.sided")

```

## For exp2: 
```{r}
library(pROC)
library(readr)
setwd("Z:/Cristina/Section3/paper_notes_section3_MODIFIED")
exp2_pooled_pred_train <- read_csv("datasets/exp2_pooled_pred_train.csv")
exp2_pooled_pred_val <- read_csv("datasets/exp2_pooled_pred_val.csv")
exp2_pred_test <- read_csv("datasets/exp2_pred_test.csv")

# Create 6 objects: Tr_exp2, ciobj_Tr_exp2, Val_exp2, ciobj_val_exp2, Te_exp2, ciobj_Te_exp2
# plot ROC curve
################
library(pROC)
par(mfrow = c(1, 3))
par(cex = 0.6)
par(mar = c(3, 3, 0, 0), oma = c(0.51, 0.51, 0.51, 0.51))
n=12
colors = rainbow(n, s = 1, v = 1, start = 0, end = max(1, n - 1)/n, alpha = 1)
# plot
Tr_exp2 <- plot.roc(exp2_pooled_pred_train$labels, exp2_pooled_pred_train$probC, col=colors[1], lty=1)
ciobj_Tr_exp2 <- ci.se(Tr_exp2, specificities=seq(0, 1, 0.05)) 
plot(ciobj_Tr_exp2, type="shape", col="grey") # plot as a grey shape
par(new=TRUE)
plot.roc(exp2_pooled_pred_train$labels, exp2_pooled_pred_train$probC, col=colors[1], lty=1, main="ROC for cvTrain")
legend("bottomright", 
       legend = c(paste0("cvTrain: AUC=", format(Tr_exp2$auc,digits=3, format="f"))), 
       col = colors[1],lwd = 2, lty = c(1))

# plot
Val_exp2 <- plot.roc(exp2_pooled_pred_val$labels, exp2_pooled_pred_val$probC, col=colors[1], lty=1)
ciobj_val_exp2 <- ci.se(Val_exp2, specificities=seq(0, 1, 0.05)) 
plot(ciobj_val_exp2, type="shape", col="grey") # plot as a blue shape
par(new=TRUE)
plot.roc(exp2_pooled_pred_val$labels, exp2_pooled_pred_val$probC, col=colors[10], lty=1, main="ROC for cvValidation")
legend("bottomright", 
       legend = c(paste0("cvValidation: AUC=", format(Val_exp2$auc, digits=3, format="f"))), 
       col = colors[10],lwd = 2, lty = c(1))
# plot
Te_exp2 <- plot.roc(exp2_pred_test$labels, exp2_pred_test$probC, col=colors[8], lty=1)
ciobj_Te_exp2 <- ci.se(Te_exp2, specificities=seq(0, 1, 0.05)) 
plot(ciobj_Te_exp2, type="shape", col="grey") # plot as a grey shape
par(new=TRUE)
plot.roc(exp2_pred_test$labels, exp2_pred_test$probC, col=colors[8], lty=1, main="ROC for held-out Test")
legend("bottomright", 
       legend = c(paste0("Test: AUC=", formatC(Te_exp2$auc,digits=2, format="f"))), 
       col = colors[8],lwd = 2, lty = c(1))

```

* Do we significantly overtrain?

```{r exp2-overtrain, echo=TRUE}
roc.test(Tr_exp2, Val_exp2, method="bootstrap",boot.stratified=TRUE, alternative="greater")

roc.test(Tr_exp2, Te_exp2, method="bootstrap",boot.stratified=TRUE, alternative="greater")

roc.test(Val_exp2, Te_exp2, method="bootstrap",boot.stratified=TRUE, alternative="two.sided")
```

## For exp3: 
```{r}
library(pROC)
library(readr)
setwd("Z:/Cristina/Section3/paper_notes_section3_MODIFIED")
exp3_pooled_pred_train <- read_csv("datasets/exp3_pooled_pred_train.csv")
exp3_pooled_pred_val <- read_csv("datasets/exp3_pooled_pred_val.csv")
exp3_pred_test <- read_csv("datasets/exp3_pred_test.csv")

# Create 6 objects: Tr_exp3, ciobj_Tr_exp3, Val_exp3, ciobj_val_exp3, Te_exp3, ciobj_Te_exp3
# plot ROC curve
################
library(pROC)
par(mfrow = c(1, 3))
par(cex = 0.6)
par(mar = c(3, 3, 0, 0), oma = c(0.51, 0.51, 0.51, 0.51))
n=12
colors = rainbow(n, s = 1, v = 1, start = 0, end = max(1, n - 1)/n, alpha = 1)
# plot
Tr_exp3 <- plot.roc(exp3_pooled_pred_train$labels, exp3_pooled_pred_train$probC, col=colors[1], lty=1)
ciobj_Tr_exp3 <- ci.se(Tr_exp3, specificities=seq(0, 1, 0.05)) 
plot(ciobj_Tr_exp3, type="shape", col="grey") # plot as a grey shape
par(new=TRUE)
plot.roc(exp3_pooled_pred_train$labels, exp3_pooled_pred_train$probC, col=colors[1], lty=1, main="ROC for cvTrain")
legend("bottomright", 
       legend = c(paste0("cvTrain: AUC=", format(Tr_exp3$auc,digits=3, format="f"))), 
       col = colors[1],lwd = 2, lty = c(1))

# plot
Val_exp3 <- plot.roc(exp3_pooled_pred_val$labels, exp3_pooled_pred_val$probC, col=colors[1], lty=1)
ciobj_val_exp3 <- ci.se(Val_exp3, specificities=seq(0, 1, 0.05)) 
plot(ciobj_val_exp3, type="shape", col="grey") # plot as a blue shape
par(new=TRUE)
plot.roc(exp3_pooled_pred_val$labels, exp3_pooled_pred_val$probC, col=colors[10], lty=1, main="ROC for cvValidation")
legend("bottomright", 
       legend = c(paste0("cvValidation: AUC=", format(Val_exp3$auc, digits=3, format="f"))), 
       col = colors[10],lwd = 2, lty = c(1))
# plot
Te_exp3 <- plot.roc(exp3_pred_test$labels, exp3_pred_test$probC, col=colors[8], lty=1)
ciobj_Te_exp3 <- ci.se(Te_exp3, specificities=seq(0, 1, 0.05)) 
plot(ciobj_Te_exp3, type="shape", col="grey") # plot as a grey shape
par(new=TRUE)
plot.roc(exp3_pred_test$labels, exp3_pred_test$probC, col=colors[8], lty=1, main="ROC for held-out Test")
legend("bottomright", 
       legend = c(paste0("Test: AUC=", formatC(Te_exp3$auc,digits=2, format="f"))), 
       col = colors[8],lwd = 2, lty = c(1))

```

* Do we significantly overtrain?

```{r exp3-overtrain, echo=TRUE}
roc.test(Tr_exp3, Val_exp3, method="bootstrap",boot.stratified=TRUE, alternative="greater")

roc.test(Tr_exp3, Te_exp3, method="bootstrap",boot.stratified=TRUE, alternative="greater")

roc.test(Val_exp3, Te_exp3, method="bootstrap",boot.stratified=TRUE, alternative="two.sided")
```

### Significance Tests for comparing two or more ROC areas:


```{r roc-comp1, echo=TRUE}
Val_exp1
Val_exp3
roc.test(Val_exp1, Val_exp3, method="bootstrap",boot.stratified=TRUE, alternative="two.sided")

Te_exp1
Te_exp3
roc.test(Te_exp1, Te_exp3, method="bootstrap",boot.stratified=TRUE, alternative="two.sided")

```

```{r roc-comp2, echo=TRUE}
Val_exp2
Val_exp3
roc.test(Val_exp2, Val_exp3, method="bootstrap",boot.stratified=TRUE, alternative="two.sided")

Te_exp2
Te_exp3
roc.test(Te_exp2, Te_exp3, method="bootstrap",boot.stratified=TRUE, alternative="two.sided")

```